---
output: 
  pdf_document: 
    highlight: haddock
    fig_height: 5
fontsize: 11pt
linestretch: 1.5
---



\begin{titlepage}
	
	\begin{center}
		
		\large{FEDERAL UNIVERSITY OF MINAS GERAIS}\\
        \vspace{2pt}
		\large{FACULTY OF ECONOMIC SCIENCES}\\
		    \vspace{2pt}
		\large{CENTER FOR DEVELOPMENT AND REGIONAL PLANNING}\\
        \vspace{8cm}
    
  \textbf{\huge{Life expectancy forecast in São Paulo city, Brazil, in the context of covid-19}}\\
      \vspace{25pt}
    \textbf{\Large{Men}}\\
        \vspace{6cm}
		    \Large{Maria Laura Lopes Miranda}\\
		    \vspace{25pt}
		
		\large{June}\\
		\large{2022}\\
	
	\end{center}

\end{titlepage}



```{r setup, include=FALSE}
rm(list = ls())
knitr::opts_chunk$set(echo = F, message = F ,warning = F, fig.align = 'center')
options(scipen = 999, OutDec = ',', digits = 6)

# Packages
library(ggplot2)
library(readxl)
library(dplyr)
library(reshape2)
library(forecast)
library(fields)
library(demography)
library(tidyverse)

```


```{r best practice}
# Life table
source("LifetableMX.R")

# Best-Practice
aux_bp_all <- read.csv('bple.csv')[90:269,] # Filter between 1840 and 2019
reg_f <- lm(e0_Fem~Year, data=aux_bp_all) # Obtain Oeppen-Vaupel curve for Fem
reg_m <- lm(e0_Masc~Year, data=aux_bp_all) # Obtain Oeppen-Vaupel curve for Mal

```



\begin{center}

\huge{Lee-Carter(1992)}

\end{center}



\subsection{Without 2020, 2021, and 2022}

```{r lc.s1}

# Data sets
E_start <- read_xlsx('exposure_5yeargroup.xlsx')[,-(1:2)] # Exposures
LMX_start <- read_xlsx('nmx.xlsx')[,-(1:2)] # Death rates
Y_start <- read_xlsx('deaths.xlsx')[,-(1:2)] # Population

# For each starting point from 1920 to 1995 (76 different combinations)
# For the forecast period 2023 - 2050 (28 different years)
# In this case it is also forecasted 2020, 2021 and 2022 but we don't include them

lc_s1 <- c()

# For using in scenarios 5 and 6
lc_LMX_2023 <- c()
lc_LMX_2024 <- c()
  
for (w in 1:76){

  # Selecting only men and excluding 2020/21/22
  E <- as.matrix(E_start[1:18,w:(length(E_start)-3)])
  LMX <- as.matrix(log(LMX_start[1:18,w:(length(LMX_start)-3)]))
  Y <- as.matrix(Y_start[1:18,w:(length(Y_start)-3)])
  
  a <- as.integer(c(0,1,seq(5,80,5))) # Age groups
  t1 <- as.integer(seq(1919+w,2019,1)) # Baseline period
  tF <- as.integer(seq(2020,2050,1)) # Forecasted horizon
  
  n1 <- length(t1)
  nF <- length(tF)
  
  Alpha <- apply(LMX,1,mean) # Mean until 2019
  
  M.tilde <- LMX - Alpha
  
  svdM <- svd(M.tilde,nu=1,nv=1)
  
  Beta <- c(svdM$u)
  Kappa1 <- c(svdM$v)
  
  sum.Beta <- sum(Beta)
  Beta <- Beta/sum.Beta
  
  Kappa1 <- Kappa1 - mean(Kappa1)
  Kappa1 <- Kappa1*sum.Beta
  
  Kappa <- rep(NA,n1)
  
  # Adjusting kt for the deaths
  koptim <- function(par,Alpha,Beta,sum.dx,Exp){
    Kappa <- par[1]
    lmx.lc <- Alpha+Beta*Kappa
    z.lc <- exp(lmx.lc)*Exp
    sum.z.lc <- sum(z.lc)
    diff.lc <- abs(sum.dx-sum.z.lc)
    return(diff.lc)
  }
  
  for (i in 1:n1){
    KappaSecStep <- optimize(f=koptim, interval=c(-150,150), Alpha=Alpha,
                             Beta=Beta, sum.dx=sum(Y[,i]), Exp=E[,i])
    Kappa[i] <- KappaSecStep$minimum
  }
  
  Kts <- ts(c(Kappa), start = t1[1])
  
  modK <- Arima(Kts, order=c(0,1,0), include.drift=TRUE)
  
  nS <- 100
  SIMk <- c()
  SIMnmx <- c()
  for(s in 1:nS){
    s <- 1
    ## generate simulation with bootsrapping
    kappa.sim <- simulate(modK, nsim=nF,future=TRUE, bootstrap=TRUE)
    kappa.matrix <- matrix(kappa.sim)
    
    OneF <- rep(1,nF)
    # Calculate nmx
    nmx <- Alpha%*%t(OneF) + Beta%*%t(kappa.matrix)
    # Calculate le by horizon
    e0 <- apply(exp(nmx),2,e0.mx,x=a,sex="M")
    
    SIMk <- rbind(SIMk, e0) # Saving the e0
    SIMnmx <- rbind(SIMnmx, kappa.sim) # Saving the kappas
  }
  
  # Saving kappas 1 column per forecasted year, one row per sim and starting year
  lc_s1 <- rbind(lc_s1, SIMk[,4:nF]) # Exclude 2020/21/22 from the horizon
  
  ## 2023 and 2024
  lc_s1_2023 <- as.matrix(SIMnmx[,4])
  lc_s1_2024 <- as.matrix(SIMnmx[,5])
  
  # Taking the mean simulation at each iteration of starting point
  kappa_2023 <- median(lc_s1_2023)
  kappa_2024 <- median(lc_s1_2024)
  
  # Getting the LMX for each starting point
  lc_LMX_2023 <- rbind(lc_LMX_2023, 
                      Alpha + Beta%*%t(kappa_2023))
  lc_LMX_2024 <- rbind(lc_LMX_2024, 
                      Alpha + Beta%*%t(kappa_2024))
  
}

```




\subsection{With 2020, 2021, and 2022}


```{r lc.s2}

# Data sets
E_start <- read_xlsx('exposure_5yeargroup.xlsx')[,-(1:2)] # Exposures
LMX_start <- read_xlsx('nmx.xlsx')[,-(1:2)] # Death rates
Y_start <- read_xlsx('deaths.xlsx')[,-(1:2)] # Population

# For each starting point from 1920 to 1995 (76 different combinations)
# For the forecast period 2023 - 2050 (28 different years)

lc_s2 <- c()

for (w in 1:76){

  # Selecting only men , data from 1920 to 2022
  E <- as.matrix(E_start[1:18,w:length(E_start)])
  LMX <- as.matrix(log(LMX_start[1:18,w:length(LMX_start)]))
  Y <- as.matrix(Y_start[1:18,w:length(Y_start)])
  
  a <- as.integer(c(0,1,seq(5,80,5))) # Age groups
  t1 <- as.integer(seq(1919+w,2022,1)) # Baseline period
  tF <- as.integer(seq(2023,2050,1)) # Forecasted horizon
  
  n1 <- length(t1)
  nF <- length(tF)
  
  Alpha <- apply(LMX[,1:(ncol(LMX)-3)],1,mean) # Mean until 2019 
  
  M.tilde <- LMX - Alpha
  
  svdM <- svd(M.tilde,nu=1,nv=1)
  
  Beta <- c(svdM$u)
  Kappa1 <- c(svdM$v)
  
  sum.Beta <- sum(Beta)
  Beta <- Beta/sum.Beta
  
  Kappa1 <- Kappa1 - mean(Kappa1)
  Kappa1 <- Kappa1*sum.Beta
  
  Kappa <- rep(NA,n1)
  
  # Adjusting kt for the deaths
  koptim <- function(par,Alpha,Beta,sum.dx,Exp){
    Kappa <- par[1]
    lmx.lc <- Alpha+Beta*Kappa
    z.lc <- exp(lmx.lc)*Exp
    sum.z.lc <- sum(z.lc)
    diff.lc <- abs(sum.dx-sum.z.lc)
    return(diff.lc)
  }
  
  for (i in 1:n1){
    KappaSecStep <- optimize(f=koptim, interval=c(-150,150), Alpha=Alpha,
                             Beta=Beta, sum.dx=sum(Y[,i]), Exp=E[,i])
    Kappa[i] <- KappaSecStep$minimum
  }
  
  Kts <- ts(c(Kappa), start = t1[1])
  
  modK <- Arima(Kts, order=c(0,1,0), include.drift=TRUE)
  
  nS <- 100
  SIMk <- c()
  for(s in 1:nS){
    s <- 1
    ## generate simulation with bootsrapping
    kappa.sim <- simulate(modK, nsim=nF,future=TRUE, bootstrap=TRUE)
    kappa.matrix <- matrix(kappa.sim)
    
    OneF <- rep(1,nF)
    # Calculate nmx
    nmx <- Alpha%*%t(OneF) + Beta%*%t(kappa.matrix)
    # Calculate le by horizon
    e0 <- apply(exp(nmx),2,e0.mx,x=a,sex="M")
    
    SIMk <- rbind(SIMk, e0) # Saving the e0
  }
  
  # Saving kappas 1 column per forecasted year, one row per sim and starting year
  lc_s2 <- rbind(lc_s2, SIMk)

}
```




\subsection{With 2020, 2021, 2022, and 2023 == 2019}

```{r lc.s3}

# Data sets
E_start <- read_xlsx('exposure_5yeargroup.xlsx')[,-(1:2)] # Exposures
LMX_start <- read_xlsx('nmx.xlsx')[,-(1:2)] # Death rates
Y_start <- read_xlsx('deaths.xlsx')[,-(1:2)] # Population

# 2023 == 2019
E_start$'2023' <- c(E_start$'2019')
LMX_start$'2023' <- c(LMX_start$'2019')
Y_start$'2023' <- c(Y_start$'2019')

# For each starting point from 1920 to 1995 (76 different combinations)
# For the forecast period 2024 - 2050 (27 different years)

lc_s3 <- c()

for (w in 1:76){

  # Selecting only men, data from 1920 to 2023
  E <- as.matrix(E_start[1:18,w:length(E_start)])
  LMX <- as.matrix(log(LMX_start[1:18,w:length(LMX_start)]))
  Y <- as.matrix(Y_start[1:18,w:length(Y_start)])
  
  a <- as.integer(c(0,1,seq(5,80,5))) # Age groups
  t1 <- as.integer(seq(1919+w,2023,1)) # Baseline period
  tF <- as.integer(seq(2024,2050,1)) # Forecasted horizon
  
  n1 <- length(t1)
  nF <- length(tF)
  
  Alpha <- apply(LMX[,1:(ncol(LMX)-4)],1,mean) # Mean until 2019
  
  M.tilde <- LMX - Alpha
  
  svdM <- svd(M.tilde,nu=1,nv=1)
  
  Beta <- c(svdM$u)
  Kappa1 <- c(svdM$v)
  
  sum.Beta <- sum(Beta)
  Beta <- Beta/sum.Beta
  
  Kappa1 <- Kappa1 - mean(Kappa1)
  Kappa1 <- Kappa1*sum.Beta
  
  Kappa <- rep(NA,n1)
  
  # Adjusting kt for the deaths
  koptim <- function(par,Alpha,Beta,sum.dx,Exp){
    Kappa <- par[1]
    lmx.lc <- Alpha+Beta*Kappa
    z.lc <- exp(lmx.lc)*Exp
    sum.z.lc <- sum(z.lc)
    diff.lc <- abs(sum.dx-sum.z.lc)
    return(diff.lc)
  }
  
  for (i in 1:n1){
    KappaSecStep <- optimize(f=koptim, interval=c(-150,150), Alpha=Alpha,
                             Beta=Beta, sum.dx=sum(Y[,i]), Exp=E[,i])
    Kappa[i] <- KappaSecStep$minimum
  }
  
  Kts <- ts(c(Kappa), start = t1[1])
  
  modK <- Arima(Kts, order=c(0,1,0), include.drift=TRUE)
  
  nS <- 100
  SIMk <- c()
  for(s in 1:nS){
    s <- 1
    ## generate simulation with bootsrapping
    kappa.sim <- simulate(modK, nsim=nF,future=TRUE, bootstrap=TRUE)
    kappa.matrix <- matrix(kappa.sim)
    
    OneF <- rep(1,nF)
    # Calculate nmx
    nmx <- Alpha%*%t(OneF) + Beta%*%t(kappa.matrix)
    # Calculate le by horizon
    e0 <- apply(exp(nmx),2,e0.mx,x=a,sex="M")
    
    SIMk <- rbind(SIMk, e0) # Saving the e0
  }
  
  # Saving kappas 1 column per forecasted year, one row per sim and starting year
  lc_s3 <- rbind(lc_s3, cbind(NA, SIMk)) # NA for the year 2023

}

```




\subsection{With 2020, 2021, 2022, and 2023 == 2023 from 1st approach}


```{r lc.s4}

# Data sets
E_start <- read_xlsx('exposure_5yeargroup.xlsx')[,-(1:2)] # Exposures
LMX_start <- read_xlsx('nmx.xlsx')[,-(1:2)] # Death rates
Y_start <- read_xlsx('deaths.xlsx')[,-(1:2)] # Population

# For each starting point from 1920 to 1995 (76 different combinations)
# For the forecast period 2024 - 2050 (27 different years)

lc_s4 <- c()

for (w in 1:76){

  # 2023 == pop interpolated for both men and women
  aux_pop <- c(61307,239866,344865,335745,364462,419659,430038,428515,450675,
               448576,377244,348645,310696,269849,213364,154532,95977,106656,
               59508,231481,332120,322157,347892,415220,439690,451424,496303,
               510045,441217,411083,383356,353289,295993,229175,155581,213611) 
  
  # Taking the estimates from s1 per iteration
  aux_nmx <- exp(c(lc_LMX_2023[((w*18)-17):(w*18),], rep(1,18)))
  aux_deaths <- aux_pop*aux_nmx
  
  E_start$'2023' <- c(aux_pop)
  LMX_start$'2023' <- c(aux_nmx)
  Y_start$'2023' <- c(aux_deaths)
  
  # Selecting only men, data from 1920 to 2023
  E <- as.matrix(E_start[1:18,w:length(E_start)])
  LMX <- as.matrix(log(LMX_start[1:18,w:length(LMX_start)]))
  Y <- as.matrix(Y_start[1:18,w:length(Y_start)])
  
  a <- as.integer(c(0,1,seq(5,80,5))) # Age groups
  t1 <- as.integer(seq(1919+w,2023,1)) # Baseline period
  tF <- as.integer(seq(2024,2050,1)) # Forecasted horizon
  
  n1 <- length(t1)
  nF <- length(tF)
  
  Alpha <- apply(LMX[,1:(ncol(LMX)-4)],1,mean) # Mean until 2019
  
  M.tilde <- LMX - Alpha
  
  svdM <- svd(M.tilde,nu=1,nv=1)
  
  Beta <- c(svdM$u)
  Kappa1 <- c(svdM$v)
  
  sum.Beta <- sum(Beta)
  Beta <- Beta/sum.Beta
  
  Kappa1 <- Kappa1 - mean(Kappa1)
  Kappa1 <- Kappa1*sum.Beta
  
  Kappa <- rep(NA,n1)
  
  # Adjusting kt for the deaths
  koptim <- function(par,Alpha,Beta,sum.dx,Exp){
    Kappa <- par[1]
    lmx.lc <- Alpha+Beta*Kappa
    z.lc <- exp(lmx.lc)*Exp
    sum.z.lc <- sum(z.lc)
    diff.lc <- abs(sum.dx-sum.z.lc)
    return(diff.lc)
  }
  
  for (i in 1:n1){
    KappaSecStep <- optimize(f=koptim, interval=c(-150,150), Alpha=Alpha,
                             Beta=Beta, sum.dx=sum(Y[,i]), Exp=E[,i])
    Kappa[i] <- KappaSecStep$minimum
  }
  
  Kts <- ts(c(Kappa), start = t1[1])
  
  modK <- Arima(Kts, order=c(0,1,0), include.drift=TRUE)
  
  nS <- 100
  SIMk <- c()
  for(s in 1:nS){
    s <- 1
    ## generate simulation with bootsrapping
    kappa.sim <- simulate(modK, nsim=nF,future=TRUE, bootstrap=TRUE)
    kappa.matrix <- matrix(kappa.sim)
    
    OneF <- rep(1,nF)
    # Calculate nmx
    nmx <- Alpha%*%t(OneF) + Beta%*%t(kappa.matrix)
    # Calculate le by horizon
    e0 <- apply(exp(nmx),2,e0.mx,x=a,sex="M")
    
    SIMk <- rbind(SIMk, e0) # Saving the e0
  }
  
  # Saving kappas 1 column per forecasted year, one row per sim and starting year
  lc_s4 <- rbind(lc_s4, cbind(NA, SIMk)) # NA for the year 2023

}

```





\subsection{With 2020, 2021, 2022, 2023 == mean(2022,2024, and 2024 == 2019}

```{r lc.s5}

# Data sets
E_start <- read_xlsx('exposure_5yeargroup.xlsx')[,-(1:2)] # Exposures
LMX_start <- read_xlsx('nmx.xlsx')[,-(1:2)] # Death rates
Y_start <- read_xlsx('deaths.xlsx')[,-(1:2)] # Population

# 2023 = mean (2022, 2024==2019)
E_start$'2023' <- rowMeans(cbind(E_start$'2022', E_start$'2019'))
LMX_start$'2023' <- rowMeans(cbind(LMX_start$'2022', LMX_start$'2019'))
Y_start$'2023' <- rowMeans(cbind(Y_start$'2022', Y_start$'2019'))

# 2024 == 2019
E_start$'2024' <- c(E_start$'2019')
LMX_start$'2024' <- c(LMX_start$'2019')
Y_start$'2024' <- c(Y_start$'2019')

# For each starting point from 1920 to 1995 (76 different combinations)
# For the forecast period 2025 - 2050 (26 different years)

lc_s5 <- c()

for (w in 1:76){

  # Selecting only men, data from 1920 to 2023
  E <- as.matrix(E_start[1:18,w:length(E_start)])
  LMX <- as.matrix(log(LMX_start[1:18,w:length(LMX_start)]))
  Y <- as.matrix(Y_start[1:18,w:length(Y_start)])
  
  a <- as.integer(c(0,1,seq(5,80,5))) # Age groups
  t1 <- as.integer(seq(1919+w,2024,1)) # Baseline period
  tF <- as.integer(seq(2025,2050,1)) # Forecasted horizon
  
  n1 <- length(t1)
  nF <- length(tF)
  
  Alpha <- apply(LMX[,1:(ncol(LMX)-5)],1,mean) # Mean until 2019
  
  M.tilde <- LMX - Alpha
  
  svdM <- svd(M.tilde,nu=1,nv=1)
  
  Beta <- c(svdM$u)
  Kappa1 <- c(svdM$v)
  
  sum.Beta <- sum(Beta)
  Beta <- Beta/sum.Beta
  
  Kappa1 <- Kappa1 - mean(Kappa1)
  Kappa1 <- Kappa1*sum.Beta
  
  Kappa <- rep(NA,n1)
  
  # Adjusting kt for the deaths
  koptim <- function(par,Alpha,Beta,sum.dx,Exp){
    Kappa <- par[1]
    lmx.lc <- Alpha+Beta*Kappa
    z.lc <- exp(lmx.lc)*Exp
    sum.z.lc <- sum(z.lc)
    diff.lc <- abs(sum.dx-sum.z.lc)
    return(diff.lc)
  }
  
  for (i in 1:n1){
    KappaSecStep <- optimize(f=koptim, interval=c(-150,150), Alpha=Alpha,
                             Beta=Beta, sum.dx=sum(Y[,i]), Exp=E[,i])
    Kappa[i] <- KappaSecStep$minimum
  }
  
  Kts <- ts(c(Kappa), start = t1[1])
  
  modK <- Arima(Kts, order=c(0,1,0), include.drift=TRUE)
  
  nS <- 100
  SIMk <- c()
  for(s in 1:nS){
    s <- 1
    ## generate simulation with bootsrapping
    kappa.sim <- simulate(modK, nsim=nF,future=TRUE, bootstrap=TRUE)
    kappa.matrix <- matrix(kappa.sim)
    
    OneF <- rep(1,nF)
    # Calculate nmx
    nmx <- Alpha%*%t(OneF) + Beta%*%t(kappa.matrix)
    # Calculate le by horizon
    e0 <- apply(exp(nmx),2,e0.mx,x=a,sex="M")
    
    SIMk <- rbind(SIMk, e0) # Saving the e0
  }
  
  # Saving kappas 1 column per forecasted year, one row per sim and starting year
  lc_s5 <- rbind(lc_s5, cbind(NA, NA, SIMk)) # NA for the years 2023/2024

}

```




\subsection{With 2020, 2021, 2022, 2023 == 2019, and 2024 == 2024 from 1st approach}


```{r lc.s6}

# Data sets
E_start <- read_xlsx('exposure_5yeargroup.xlsx')[,-(1:2)] # Exposures
LMX_start <- read_xlsx('nmx.xlsx')[,-(1:2)] # Death rates
Y_start <- read_xlsx('deaths.xlsx')[,-(1:2)] # Population

# For each starting point from 1920 to 1995 (76 different combinations)
# For the forecast period 2025 - 2050 (26 different years)

lc_s6 <- c()

for (w in 1:76){

  # 2024 == pop interpolated for both men and women
  aux_pop <- c(60438,236467,341870,328906,360489,414672,423799,424694,452902,
               453881,380135,352523,316615,277903,222212,160419,98961,110709,
               58672,228227,329210,315071,342781,409158,431822,445929,498793,
               517072,444482,414790,390324,364001,308002,238295,160062,220764)  
  
  # Taking the estimates from s1 per iteration
  aux_nmx <- exp(c(lc_LMX_2024[((w*18)-17):(w*18),], rep(1,18)))
  aux_deaths <- aux_pop*aux_nmx
  
  # 2023 == mean (2022, 2024)
  E_start$'2023' <- rowMeans(cbind(E_start$'2022', aux_pop))
  LMX_start$'2023' <- rowMeans(cbind(LMX_start$'2022', aux_nmx))
  Y_start$'2023' <- rowMeans(cbind(Y_start$'2022', aux_deaths))
  
  E_start$'2024' <- c(aux_pop)
  LMX_start$'2024' <- c(aux_nmx)
  Y_start$'2024' <- c(aux_deaths)
  
  # Selecting only men, data from 1920 to 2024
  E <- as.matrix(E_start[1:18,w:length(E_start)])
  LMX <- as.matrix(log(LMX_start[1:18,w:length(LMX_start)]))
  Y <- as.matrix(Y_start[1:18,w:length(Y_start)])
  
  a <- as.integer(c(0,1,seq(5,80,5))) # Age groups
  t1 <- as.integer(seq(1919+w,2024,1)) # Baseline period
  tF <- as.integer(seq(2025,2050,1)) # Forecasted horizon
  
  n1 <- length(t1)
  nF <- length(tF)
  
  Alpha <- apply(LMX[,1:(ncol(LMX)-5)],1,mean) # Mean until 2019
  
  M.tilde <- LMX - Alpha
  
  svdM <- svd(M.tilde,nu=1,nv=1)
  
  Beta <- c(svdM$u)
  Kappa1 <- c(svdM$v)
  
  sum.Beta <- sum(Beta)
  Beta <- Beta/sum.Beta
  
  Kappa1 <- Kappa1 - mean(Kappa1)
  Kappa1 <- Kappa1*sum.Beta
  
  Kappa <- rep(NA,n1)
  
  # Adjusting kt for the deaths
  koptim <- function(par,Alpha,Beta,sum.dx,Exp){
    Kappa <- par[1]
    lmx.lc <- Alpha+Beta*Kappa
    z.lc <- exp(lmx.lc)*Exp
    sum.z.lc <- sum(z.lc)
    diff.lc <- abs(sum.dx-sum.z.lc)
    return(diff.lc)
  }
  
  for (i in 1:n1){
    KappaSecStep <- optimize(f=koptim, interval=c(-150,150), Alpha=Alpha,
                             Beta=Beta, sum.dx=sum(Y[,i]), Exp=E[,i])
    Kappa[i] <- KappaSecStep$minimum
  }
  
  Kts <- ts(c(Kappa), start = t1[1])
  
  modK <- Arima(Kts, order=c(0,1,0), include.drift=TRUE)
  
  nS <- 100
  SIMk <- c()
  for(s in 1:nS){
    s <- 1
    ## generate simulation with bootsrapping
    kappa.sim <- simulate(modK, nsim=nF,future=TRUE, bootstrap=TRUE)
    kappa.matrix <- matrix(kappa.sim)
    
    OneF <- rep(1,nF)
    # Calculate nmx
    nmx <- Alpha%*%t(OneF) + Beta%*%t(kappa.matrix)
    # Calculate le by horizon
    e0 <- apply(exp(nmx),2,e0.mx,x=a,sex="M")
    
    SIMk <- rbind(SIMk, e0) # Saving the e0
  }
  
  # Saving kappas 1 column per forecasted year, one row per sim and starting year
  lc_s6 <- rbind(lc_s6, cbind(NA, NA, SIMk)) # NA for the years 2023/2024

}

```





\newpage
\begin{center}

\huge{Lee-Miller(2001)}

\end{center}



\subsection{Without 2020, 2021, and 2022}

```{r lm.s1}

# Data sets
E_start <- read_xlsx('exposure_5yeargroup.xlsx')[,-(1:2)] # Exposures
LMX_start <- read_xlsx('nmx.xlsx')[,-(1:2)] # Death rates
Y_start <- read_xlsx('deaths.xlsx')[,-(1:2)] # Population

# For each starting point from 1950 to 1995 (46 different combinations)
# For the forecast period 2023 - 2050 (28 different years)
# In this case it is also forecasted 2020, 2021, and 2022 but we don't include them

lm_s1 <- c()

# For using in scenarios 5 and 6
lm_LMX_2023 <- c()
lm_LMX_2024 <- c()

for (w in 31:76){

  # Selecting only men and excluding 2020/21/22
  E <- as.matrix(E_start[1:18,w:(length(E_start)-3)])
  LMX <- as.matrix(log(LMX_start[1:18,w:(length(LMX_start)-3)]))
  Y <- as.matrix(Y_start[1:18,w:(length(Y_start)-3)])
  
  a <- as.integer(c(0,1,seq(5,80,5))) # Age groups
  t1 <- as.integer(seq(1919+w,2019,1)) # Baseline period
  tF <- as.integer(seq(2020,2050,1)) # Forecasted horizon
  
  n1 <- length(t1)
  nF <- length(tF)
  
  Alpha <- LMX[,ncol(LMX)] #### Instead of mean use the year of 2019
  
  M.tilde <- LMX - Alpha
  
  svdM <- svd(M.tilde,nu=1,nv=1)
  
  Beta <- c(svdM$u)
  Kappa1 <- c(svdM$v)
  
  sum.Beta <- sum(Beta)
  Beta <- Beta/sum.Beta
  
  Kappa1 <- Kappa1 - mean(Kappa1)
  Kappa1 <- Kappa1*sum.Beta
  
  Kappa <- rep(NA,n1)
  
  e0.obs.real <- apply(exp(LMX),2,e0.mx,x=a,sex="M")
  
  # Adjusting kt for the e0
  koptim <- function(par,Alpha,Beta,e0.obs,x,sex){
    Kappa <- par[1]
    lmx.lc <- Alpha+Beta*Kappa
    e0.est <- e0.mx(mx=exp(lmx.lc),x=a,sex="M") #e0 adjustment
    diff.lc <- abs(e0.obs-e0.est)
    return(diff.lc)
  }
  
  for (i in 1:n1){
    KappaSecStep <- optimize(f=koptim, interval=c(-150,150), Alpha=Alpha, 
                             Beta=Beta, e0.obs=e0.obs.real[i], x=a, sex="M")
    Kappa[i] <- KappaSecStep$minimum
  }
  
  Kts <- ts(c(Kappa), start = t1[1])
  
  modK <- Arima(Kts, order=c(0,1,0), include.drift=TRUE)
  
  nS <- 100
  SIMk <- c()
  SIMnmx <- c()
  for(s in 1:nS){
    s <- 1
    ## generate simulation with bootsrapping
    kappa.sim <- simulate(modK, nsim=nF,future=TRUE, bootstrap=TRUE)
    kappa.matrix <- matrix(kappa.sim)
    
    OneF <- rep(1,nF)
    # Calculate nmx
    nmx <- Alpha%*%t(OneF) + Beta%*%t(kappa.matrix)
    # Calculate le by horizon
    e0 <- apply(exp(nmx),2,e0.mx,x=a,sex="M")
    
    SIMk <- rbind(SIMk, e0) # Saving the e0
    SIMnmx <- rbind(SIMnmx, kappa.sim) # Saving the kappas
  }
  
  # Saving kappas 1 column per forecasted year, one row per sim and starting year
  lm_s1 <- rbind(lm_s1, SIMk[,4:nF]) # Excluding 2020/21/22
  
  ## 2023 and 2024
  lm_s1_2023 <- as.matrix(SIMnmx[,4])
  lm_s1_2024 <- as.matrix(SIMnmx[,5])
  
  # Taking the mean simulation at each iteration of starting point
  kappa_2023 <- median(lm_s1_2023)
  kappa_2024 <- median(lm_s1_2024)
  
  # Getting the LMX for each starting point
  lm_LMX_2023 <- rbind(lm_LMX_2023, 
                      Alpha + Beta%*%t(kappa_2023))
  lm_LMX_2024 <- rbind(lm_LMX_2024, 
                      Alpha + Beta%*%t(kappa_2024))
  
}

```




\subsection{With 2020, 2021, and 2022}

```{r lm.s2}
# Data sets
E_start <- read_xlsx('exposure_5yeargroup.xlsx')[,-(1:2)] # Exposures
LMX_start <- read_xlsx('nmx.xlsx')[,-(1:2)] # Death rates
Y_start <- read_xlsx('deaths.xlsx')[,-(1:2)] # Population

# For each starting point from 1950 to 1995 (46 different combinations)
# For the forecast period 2023 - 2050 (28 different years)

lm_s2 <- c()

for (w in 31:76){

  # Selecting only men, data from 1920 to 2021
  E <- as.matrix(E_start[1:18,w:length(E_start)])
  LMX <- as.matrix(log(LMX_start[1:18,w:length(LMX_start)]))
  Y <- as.matrix(Y_start[1:18,w:length(Y_start)])
  
  a <- as.integer(c(0,1,seq(5,80,5))) # Age groups
  t1 <- as.integer(seq(1919+w,2022,1)) # Baseline period
  tF <- as.integer(seq(2023,2050,1)) # Forecasted horizon
  
  n1 <- length(t1)
  nF <- length(tF)
  
  Alpha <- LMX[,ncol(LMX)-3] #### Instead of mean use the year of 2019
  
  M.tilde <- LMX - Alpha
  
  svdM <- svd(M.tilde,nu=1,nv=1)
  
  Beta <- c(svdM$u)
  Kappa1 <- c(svdM$v)
  
  sum.Beta <- sum(Beta)
  Beta <- Beta/sum.Beta
  
  Kappa1 <- Kappa1 - mean(Kappa1)
  Kappa1 <- Kappa1*sum.Beta
  
  Kappa <- rep(NA,n1)
  
  e0.obs.real <- apply(exp(LMX),2,e0.mx,x=a,sex="M")
  
  # Adjusting kt for the e0
  koptim <- function(par,Alpha,Beta,e0.obs,x,sex){
    Kappa <- par[1]
    lmx.lc <- Alpha+Beta*Kappa
    e0.est <- e0.mx(mx=exp(lmx.lc),x=a,sex="M") #e0 adjustment
    diff.lc <- abs(e0.obs-e0.est)
    return(diff.lc)
  }
  
  for (i in 1:n1){
    KappaSecStep <- optimize(f=koptim, interval=c(-150,150), Alpha=Alpha,
                             Beta=Beta, e0.obs=e0.obs.real[i], x=a, sex="M")
    Kappa[i] <- KappaSecStep$minimum
  }
  
  Kts <- ts(c(Kappa), start = t1[1])
  
  modK <- Arima(Kts, order=c(0,1,0), include.drift=TRUE)

  nS <- 100
  SIMk <- c()
  for(s in 1:nS){
    s <- 1
    ## generate simulation with bootsrapping
    kappa.sim <- simulate(modK, nsim=nF,future=TRUE, bootstrap=TRUE)
    kappa.matrix <- matrix(kappa.sim)
    
    OneF <- rep(1,nF)
    # Calculate nmx
    nmx <- Alpha%*%t(OneF) + Beta%*%t(kappa.matrix)
    # Calculate le by horizon
    e0 <- apply(exp(nmx),2,e0.mx,x=a,sex="M")
    
    SIMk <- rbind(SIMk, e0) # Saving the e0
  }
  
  # Saving kappas 1 column per forecasted year, one row per sim and starting year
  lm_s2 <- rbind(lm_s2, SIMk)

}

```




\subsection{With 2020, 2021, 2022, and 2023 == 2019}

```{r lm.s3}

# Data sets
E_start <- read_xlsx('exposure_5yeargroup.xlsx')[,-(1:2)] # Exposures
LMX_start <- read_xlsx('nmx.xlsx')[,-(1:2)] # Death rates
Y_start <- read_xlsx('deaths.xlsx')[,-(1:2)] # Population

# 2023 == 2019
E_start$'2023' <- c(E_start$'2019')
LMX_start$'2023' <- c(LMX_start$'2019')
Y_start$'2023' <- c(Y_start$'2019')

# For each starting point from 1950 to 1995 (46 different combinations)
# For the forecast period 2024 - 2050 (27 different years)

lm_s3 <- c()

for (w in 31:76){

  # Selecting only men, data from 1920 to 2023
  E <- as.matrix(E_start[1:18,w:length(E_start)])
  LMX <- as.matrix(log(LMX_start[1:18,w:length(LMX_start)]))
  Y <- as.matrix(Y_start[1:18,w:length(Y_start)])
  
  a <- as.integer(c(0,1,seq(5,80,5))) # Age groups
  t1 <- as.integer(seq(1919+w,2023,1)) # Baseline period
  tF <- as.integer(seq(2024,2050,1)) # Forecasted horizon
  
  n1 <- length(t1)
  nF <- length(tF)
  
  Alpha <- LMX[,ncol(LMX)-4] #### Instead of mean use the year of 2019
  
  M.tilde <- LMX - Alpha
  
  svdM <- svd(M.tilde,nu=1,nv=1)
  
  Beta <- c(svdM$u)
  Kappa1 <- c(svdM$v)
  
  sum.Beta <- sum(Beta)
  Beta <- Beta/sum.Beta
  
  Kappa1 <- Kappa1 - mean(Kappa1)
  Kappa1 <- Kappa1*sum.Beta
  
  Kappa <- rep(NA,n1)
  
  e0.obs.real <- apply(exp(LMX),2,e0.mx,x=a,sex="M")
  
  # Adjusting kt for the e0
  koptim <- function(par,Alpha,Beta,e0.obs,x,sex){
    Kappa <- par[1]
    lmx.lc <- Alpha+Beta*Kappa
    e0.est <- e0.mx(mx=exp(lmx.lc),x=a,sex="M") #e0 adjustment
    diff.lc <- abs(e0.obs-e0.est)
    return(diff.lc)
  }
  
  for (i in 1:n1){
    KappaSecStep <- optimize(f=koptim, interval=c(-150,150), Alpha=Alpha,
                             Beta=Beta, e0.obs=e0.obs.real[i], x=a, sex="M")
    Kappa[i] <- KappaSecStep$minimum
  }
  
  Kts <- ts(c(Kappa), start = t1[1])
  
  modK <- Arima(Kts, order=c(0,1,0), include.drift=TRUE)

  nS <- 100
  SIMk <- c()
  for(s in 1:nS){
    s <- 1
    ## generate simulation with bootsrapping
    kappa.sim <- simulate(modK, nsim=nF,future=TRUE, bootstrap=TRUE)
    kappa.matrix <- matrix(kappa.sim)
    
    OneF <- rep(1,nF)
    # Calculate nmx
    nmx <- Alpha%*%t(OneF) + Beta%*%t(kappa.matrix)
    # Calculate le by horizon
    e0 <- apply(exp(nmx),2,e0.mx,x=a,sex="M")
    
    SIMk <- rbind(SIMk, e0) # Saving the e0
  }
  
  # Saving kappas 1 column per forecasted year, one row per sim and starting year
  lm_s3 <- rbind(lm_s3, cbind(NA, SIMk)) # NA for the year 2023

}

```




\subsection{With 2020, 2021, 2022, and 2023 == 2023 from 1st approach}

```{r lm.s4}

# Data sets
E_start <- read_xlsx('exposure_5yeargroup.xlsx')[,-(1:2)] # Exposures
LMX_start <- read_xlsx('nmx.xlsx')[,-(1:2)] # Death rates
Y_start <- read_xlsx('deaths.xlsx')[,-(1:2)] # Population

# For each starting point from 1950 to 1995 (46 different combinations)
# For the forecast period 2024 - 2050 (27 different years)

lm_s4 <- c()

for (w in 31:76){
  
  # 2023 == pop interpolated for both men and women
  aux_pop <- c(61307,239866,344865,335745,364462,419659,430038,428515,450675,
               448576,377244,348645,310696,269849,213364,154532,95977,106656,
               59508,231481,332120,322157,347892,415220,439690,451424,496303,
               510045,441217,411083,383356,353289,295993,229175,155581,213611)
  
  # Taking the estimates from s1 per iteration
  aux_nmx <- exp(c(lm_LMX_2023[(((w-30)*18)-17):((w-30)*18),], rep(1,18)))
  aux_deaths <- aux_pop*aux_nmx
  
  E_start$'2023' <- c(aux_pop)
  LMX_start$'2023' <- c(aux_nmx)
  Y_start$'2023' <- c(aux_deaths)
  
  # Selecting only men, data from 1920 to 2023
  E <- as.matrix(E_start[1:18,w:length(E_start)])
  LMX <- as.matrix(log(LMX_start[1:18,w:length(LMX_start)]))
  Y <- as.matrix(Y_start[1:18,w:length(Y_start)])
  
  a <- as.integer(c(0,1,seq(5,80,5)))
  t1 <- as.integer(seq(1919+w,2023,1))
  tF <- as.integer(seq(2024,2050,1))
  
  n1 <- length(t1)
  nF <- length(tF)
  
  Alpha <- LMX[,ncol(LMX)-4] #### Instead of mean use the year of 2019
  
  M.tilde <- LMX - Alpha
  
  svdM <- svd(M.tilde,nu=1,nv=1)
  
  Beta <- c(svdM$u)
  Kappa1 <- c(svdM$v)
  
  sum.Beta <- sum(Beta)
  Beta <- Beta/sum.Beta
  
  Kappa1 <- Kappa1 - mean(Kappa1)
  Kappa1 <- Kappa1*sum.Beta
  
  Kappa <- rep(NA,n1)
  
  e0.obs.real <- apply(exp(LMX),2,e0.mx,x=a,sex="M")
  
  # Adjusting kt for the e0
  koptim <- function(par,Alpha,Beta,e0.obs,x,sex){
    Kappa <- par[1]
    lmx.lc <- Alpha+Beta*Kappa
    e0.est <- e0.mx(mx=exp(lmx.lc),x=a,sex="M") #e0 adjustment
    diff.lc <- abs(e0.obs-e0.est)
    return(diff.lc)
  }
  
  for (i in 1:n1){
    KappaSecStep <- optimize(f=koptim, interval=c(-150,150), Alpha=Alpha,
                             Beta=Beta, e0.obs=e0.obs.real[i], x=a, sex="M")
    Kappa[i] <- KappaSecStep$minimum
  }
  
  Kts <- ts(c(Kappa), start = t1[1])
  
  modK <- Arima(Kts, order=c(0,1,0), include.drift=TRUE)

  nS <- 100
  SIMk <- c()
  for(s in 1:nS){
    s <- 1
    ## generate simulation with bootsrapping
    kappa.sim <- simulate(modK, nsim=nF,future=TRUE, bootstrap=TRUE)
    kappa.matrix <- matrix(kappa.sim)
    
    OneF <- rep(1,nF)
    # Calculate nmx
    nmx <- Alpha%*%t(OneF) + Beta%*%t(kappa.matrix)
    # Calculate le by horizon
    e0 <- apply(exp(nmx),2,e0.mx,x=a,sex="M")
    
    SIMk <- rbind(SIMk, e0) # Saving the e0
  }
  
  # Saving kappas 1 column per forecasted year, one row per sim and starting year
  lm_s4 <- rbind(lm_s4, cbind(NA, SIMk)) # NA for the year 2023

}

```




\subsection{With 2020, 2021, 2022, 2023 == mean(2022, 2024), and 2024 == 2019}

```{r lm.s5}

# Data sets
E_start <- read_xlsx('exposure_5yeargroup.xlsx')[,-(1:2)] # Exposures
LMX_start <- read_xlsx('nmx.xlsx')[,-(1:2)] # Death rates
Y_start <- read_xlsx('deaths.xlsx')[,-(1:2)] # Population

# 2023 = mean (2022, 2024==2019)
E_start$'2023' <- rowMeans(cbind(E_start$'2022', E_start$'2019'))
LMX_start$'2023' <- rowMeans(cbind(LMX_start$'2022', LMX_start$'2019'))
Y_start$'2023' <- rowMeans(cbind(Y_start$'2022', Y_start$'2019'))

# 2024 == 2019
E_start$'2024' <- c(E_start$'2019')
LMX_start$'2024' <- c(LMX_start$'2019')
Y_start$'2024' <- c(Y_start$'2019')

# For each starting point from 1950 to 1995 (46 different combinations)
# For the forecast period 2025 - 2050 (26 different years)

lm_s5 <- c()

for (w in 31:76){

  # Selecting only men, data from 1920 to 2024
  E <- as.matrix(E_start[1:18,w:length(E_start)])
  LMX <- as.matrix(log(LMX_start[1:18,w:length(LMX_start)]))
  Y <- as.matrix(Y_start[1:18,w:length(Y_start)])
  
  a <- as.integer(c(0,1,seq(5,80,5))) # Age groups
  t1 <- as.integer(seq(1919+w,2024,1)) # Baseline period
  tF <- as.integer(seq(2025,2050,1)) # Forecasted horizon
  
  n1 <- length(t1)
  nF <- length(tF)
  
  Alpha <- LMX[,ncol(LMX)-5] #### Instead of mean use the year of 2019
  
  M.tilde <- LMX - Alpha
  
  svdM <- svd(M.tilde,nu=1,nv=1)
  
  Beta <- c(svdM$u)
  Kappa1 <- c(svdM$v)
  
  sum.Beta <- sum(Beta)
  Beta <- Beta/sum.Beta
  
  Kappa1 <- Kappa1 - mean(Kappa1)
  Kappa1 <- Kappa1*sum.Beta
  
  Kappa <- rep(NA,n1)
  
  e0.obs.real <- apply(exp(LMX),2,e0.mx,x=a,sex="M")
  
  # Adjusting kt for the e0
  koptim <- function(par,Alpha,Beta,e0.obs,x,sex){
    Kappa <- par[1]
    lmx.lc <- Alpha+Beta*Kappa
    e0.est <- e0.mx(mx=exp(lmx.lc),x=a,sex="M") #e0 adjustment
    diff.lc <- abs(e0.obs-e0.est)
    return(diff.lc)
  }
  
  for (i in 1:n1){
    KappaSecStep <- optimize(f=koptim, interval=c(-150,150), Alpha=Alpha,
                             Beta=Beta, e0.obs=e0.obs.real[i], x=a, sex="M")
    Kappa[i] <- KappaSecStep$minimum
  }
  
  Kts <- ts(c(Kappa), start = t1[1])
  
  modK <- Arima(Kts, order=c(0,1,0), include.drift=TRUE)

  nS <- 100
  SIMk <- c()
  for(s in 1:nS){
    s <- 1
    ## generate simulation with bootsrapping
    kappa.sim <- simulate(modK, nsim=nF,future=TRUE, bootstrap=TRUE)
    kappa.matrix <- matrix(kappa.sim)
    
    OneF <- rep(1,nF)
    # Calculate nmx
    nmx <- Alpha%*%t(OneF) + Beta%*%t(kappa.matrix)
    # Calculate le by horizon
    e0 <- apply(exp(nmx),2,e0.mx,x=a,sex="M")
    
    SIMk <- rbind(SIMk, e0) # Saving the e0
  }
  
  # Saving kappas 1 column per forecasted year, one row per sim and starting year
  lm_s5 <- rbind(lm_s5, cbind(NA, NA, SIMk)) # NA for the year 2023/2024

}

```




\subsection{With 2020, 2021, 2022, 2023 == mean(2022, 2024), and 2024 == 2024 from 1st approach}

```{r lm.s6}

# Data sets
E_start <- read_xlsx('exposure_5yeargroup.xlsx')[,-(1:2)] # Exposures
LMX_start <- read_xlsx('nmx.xlsx')[,-(1:2)] # Death rates
Y_start <- read_xlsx('deaths.xlsx')[,-(1:2)] # Population

# For each starting point from 1950 to 1995 (46 different combinations)
# For the forecast period 2023 - 2050 (28 different years)

lm_s6 <- c()

for (w in 31:76){
  
  # 2024 == pop interpolated for both men and women
  aux_pop <- c(60438,236467,341870,328906,360489,414672,423799,424694,452902,
               453881,380135,352523,316615,277903,222212,160419,98961,110709,
               58672,228227,329210,315071,342781,409158,431822,445929,498793,
               517072,444482,414790,390324,364001,308002,238295,160062,220764) 
  
  # Taking the estimates from s1 per iteration
  aux_nmx <- exp(c(lm_LMX_2024[(((w-30)*18)-17):((w-30)*18),], rep(1,18)))
  aux_deaths <- aux_pop*aux_nmx
  
  # 2023 == mean (2022, 2024)
  E_start$'2023' <- rowMeans(cbind(E_start$'2022', aux_pop))
  LMX_start$'2023' <- rowMeans(cbind(LMX_start$'2022', aux_nmx))
  Y_start$'2023' <- rowMeans(cbind(Y_start$'2022', aux_deaths))
  
  E_start$'2024' <- c(aux_pop)
  LMX_start$'2024' <- c(aux_nmx)
  Y_start$'2024' <- c(aux_deaths)
  
  # Selecting only men, data from 1920 to 2024
  E <- as.matrix(E_start[1:18,w:length(E_start)])
  LMX <- as.matrix(log(LMX_start[1:18,w:length(LMX_start)]))
  Y <- as.matrix(Y_start[1:18,w:length(Y_start)])
  
  a <- as.integer(c(0,1,seq(5,80,5)))
  t1 <- as.integer(seq(1919+w,2024,1))
  tF <- as.integer(seq(2025,2050,1))
  
  n1 <- length(t1)
  nF <- length(tF)
  
  Alpha <- LMX[,ncol(LMX)-5] #### Instead of mean use the year of 2019
  
  M.tilde <- LMX - Alpha
  
  svdM <- svd(M.tilde,nu=1,nv=1)
  
  Beta <- c(svdM$u)
  Kappa1 <- c(svdM$v)
  
  sum.Beta <- sum(Beta)
  Beta <- Beta/sum.Beta
  
  Kappa1 <- Kappa1 - mean(Kappa1)
  Kappa1 <- Kappa1*sum.Beta
  
  Kappa <- rep(NA,n1)
  
  e0.obs.real <- apply(exp(LMX),2,e0.mx,x=a,sex="M")
  
  # Adjusting kt for the e0
  koptim <- function(par,Alpha,Beta,e0.obs,x,sex){
    Kappa <- par[1]
    lmx.lc <- Alpha+Beta*Kappa
    e0.est <- e0.mx(mx=exp(lmx.lc),x=a,sex="M") #e0 adjustment
    diff.lc <- abs(e0.obs-e0.est)
    return(diff.lc)
  }
  
  for (i in 1:n1){
    KappaSecStep <- optimize(f=koptim, interval=c(-150,150), Alpha=Alpha,
                             Beta=Beta, e0.obs=e0.obs.real[i], x=a, sex="M")
    Kappa[i] <- KappaSecStep$minimum
  }
  
  Kts <- ts(c(Kappa), start = t1[1])
  
  modK <- Arima(Kts, order=c(0,1,0), include.drift=TRUE)

  nS <- 100
  SIMk <- c()
  for(s in 1:nS){
    s <- 1
    ## generate simulation with bootsrapping
    kappa.sim <- simulate(modK, nsim=nF,future=TRUE, bootstrap=TRUE)
    kappa.matrix <- matrix(kappa.sim)
    
    OneF <- rep(1,nF)
    # Calculate nmx
    nmx <- Alpha%*%t(OneF) + Beta%*%t(kappa.matrix)
    # Calculate le by horizon
    e0 <- apply(exp(nmx),2,e0.mx,x=a,sex="M")
    
    SIMk <- rbind(SIMk, e0) # Saving the e0
  }
  
  # Saving kappas 1 column per forecasted year, one row per sim and starting year
  lm_s6 <- rbind(lm_s6, cbind(NA, NA, SIMk)) # NA for the years 2023/2024

}

```




\newpage
\begin{center}

\huge{Summary}

\Large{Lee-Carter(1992) and Lee-Miller(2001)}

\end{center}


```{r summary}

comb <- rbind(lc_s1, lc_s2, lc_s3, lc_s4, lc_s5, lc_s6,
              lm_s1, lm_s2, lm_s3, lm_s4, lm_s5, lm_s6)

colnames(comb) <- seq(2023, 2050)

## define PI level
lev <- 95
lev.p <- lev/100

# Conf interval for kappas
e0.low <- apply(comb,2,quantile,prob=(1-lev.p)/2, na.rm=T)
e0.upp <- apply(comb,2,quantile,prob=1-(1-lev.p)/2, na.rm=T)
e0.med <- apply(comb,2,median, na.rm=T)
e0.max <- apply(comb,2,max, na.rm=T)
e0.min <- apply(comb,2,min, na.rm=T)


aux <- data.frame(year = seq(2023, 2050),
                  e0.med,
                  e0.low,
                  e0.upp)

aux2 <- data.frame(year = seq(2023, 2050),
                  round(e0.med,2),
                  round(e0.low,2),
                  round(e0.upp,2))


write.table(aux2, 'summary_male.txt')

# Getting estimates for 1920-2022
LMX_start <- read_xlsx('nmx.xlsx') # Death rates
LMX <- as.matrix(log(LMX_start[1:18,3:length(LMX_start)]))

aux_obs <- data.frame(year = seq(1920, 2022),
                      e0.med = apply(exp(LMX),2,e0.mx,x=a,sex="M"))
  

reg_m <- lm(e0_Masc~Year, data=aux_bp_all) # Obtain Oeppen-Vaupel curve for Male

# Plotting the results
ggplot()+
  geom_point(data = aux_obs,
             aes(x = year, y = e0.med, color = "Observed"), lwd = 1)+
  geom_line(data = aux, 
             aes(x = year, y = e0.med, 
                 color = "Median estimate"), lwd = 1)+
  #geom_line(data = aux, 
  #           aes(x = year, y = e0.low,
  #               color = "Forecasted considering all scenarios"), 
  #          linetype = "dashed", lwd = 1, show.legend = F)+
  #geom_line(data = aux, 
  #           aes(x = year, y = e0.upp,
  #               color = "Forecasted considering all scenarios"), 
  #          linetype = "dashed", lwd = 1, show.legend = F)+
  geom_ribbon(data = aux, 
             aes(x = year, y = e0.upp,ymin=e0.low, ymax=e0.upp, fill = "95% PI"), alpha=0.2)+
  # Best-Practice
  geom_abline(aes(intercept = reg_m$coefficients[1], slope = reg_m$coefficients[2],
                  color="Best-Practice"), lwd = 1, show.legend = F)+
  theme_minimal()+
  scale_color_manual(values = c("Observed"="grey40",
                                "Median estimate"="orange",
                                "Best-Practice"="steelblue3"))+
  scale_fill_manual("",values="grey12")+
  labs(x="Year", y=""~"e"[0]~"", color = '')+
  theme(legend.position = "bottom",
        text = element_text(size = 15),
        axis.title.x = element_text(vjust = -1),
        axis.line = element_line(colour = "grey30"))+
  scale_x_continuous(breaks = scales::pretty_breaks(n = 15))+
  scale_y_continuous(limits = c(35, 97.5), n.breaks = 15)

ggsave(path = "plots/",filename = "figure_fore_male.pdf", width = 9, height = 5)
ggsave(path = "plots/",filename = "figure_fore_male.png", width = 9, height = 5)

```











